A novell interest management algorithm for distributed simulations of MMGs

### 2008-05-01 textos/dsrt/engaoi.tex
eng

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be 200 ms. If the avatar $A_i$, which just sent a state update packet to $A_j$, is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200/0.5. So, the player controlling $A_j$ will only receive an update of $A_i$ within 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequecy of $A_i$ in 50\%. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player who controls $A_j$.

@@

In order to perform the simulation of the proposed algorithm, it was first necessary to create a simulated virtual environment, with avatars present on it, because the algorithm is based on locality and view angle information. The environment consists of a two-dimensional space, which corresponds to the region managed by one of the servers. There are various avatars present, whose number varies from one simulation to another. Each one of them randomly chooses a destination point in the environment and then goes there. When it reaches its destination, it stays for a random time value, which may be zero, and then chooses a new destination to go to.

Ns-2 simulator \cite{mccanne:nsn} was used to simulate and compare the interest management algorithms described in this paper. This simulator allows the user to create code of the specific application which will be simulated. In our case, it was simulated a server which should send state updates to the client responsible for one of the avatars in the region managed by that server. Based on the location of the other avatars and on the chosen interest management algorithm, the server decided which other avatars had a relevant state to the client in question.

Through the simulations, it was obtained the server upload bandwidth usage with one client. It was not considered necessary to simulate every client connected to the server simultaneously because every one of them had the same behavior. To find the total upload bandwidth usage, we just multiply the value found in the simulation by the number of clients connected to that server. Another question is that the upload bandwidth usage by the server is much higher than download. If it receives $n$ actions, each from one of the $n$ clients connected to it, it must send, in the worst case, $O(n^2)$ state updates, because each player needs the state of the others. Therefore, it was only necessary to measure the upload traffic from the server.

@@

Some works, as \cite{yu2007nas}, \cite{kim2005tcm} and \cite{svoboda2007taa}, analyse the network traffic generated by large scale games. Based on these analysis, and adopting a conservative posture, the following parameters have been decided to be used in the simulations:

\begin{itemize}
 \item Normal update interval: 250 ms;
 \item State update packet size for one entity: 100 bytes;
 \item Duration of each simulated game session: 20 min;
 \item Virtual environment area: 750 x 750;
 \item View distance: 120;
 \item Critical distance: 40;
 \item View angle: 180°.
\end{itemize}

Several simulations were executed, in order to compare the described interest management algorithms. The number of avatars in the environment was one of the variables analyzed, to verify the scalability. The algorithms compared were the ones with circle, circle with attenuation and field of view based area of interest and the proposed algorithm, \malg. To demonstrate how each one of these reduces traffic, simulations in which no type of interest management was used - and the server sent to the client state updates of all entities in the environment - were also performed.

@@

The results were collected as follows: to measure the average upload bandwidth usage by the server, the sizes of all packets sent in the session were summed up and then divided by the session duration; to determine the maximum bandwidth usage by the server, it was measured, second by second, how many bytes had been sent and the highest value was selected.

@@

\begin{table}[ht]
\caption{Largura de banda máxima utilizada}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9400	&	8500	&	5700	&	7100	&	4700 \\
		50	&	19300	&	17000	&	10300	&	12300	&	8100 \\
		75	&	29100	&	23600	&	16600	&	17800	&	11300 \\
		100	&	38800	&	32500	&	20500	&	23000	&	15500 \\
		125	&	48600	&	37400	&	24300	&	29500	&	19700 \\
		150	&	58300	&	47400	&	29900	&	32900	&	22700 \\
		175	&	67700	&	56100	&	34300	&	32400	&	21500 \\
		200	&	77600	&	62300	&	37500	&	41200	&	28900 \\
	\hline
  \end{tabular}
\label{tab_max}
\end{table}

\begin{table}[ht]
\caption{Largura de banda utilizada em média}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9221	&	4715	&	2759	&	2534	&	1700 \\
		50	&	18826	&	9350	&	5442	&	4949	&	3303 \\
		75	&	28432	&	13963	&	8315	&	7619	&	5137 \\
		100	&	38037	&	19324	&	11029	&	9928	&	6739 \\
		125	&	47642	&	23138	&	13871	&	12434	&	8290 \\
		150	&	57247	&	29031	&	16432	&	15085	&	10062 \\
		175	&	66853	&	34697	&	19661	&	23060	&	14250 \\
		200	&	76458	& 	38600	&	23450	&	21491	&	14413 \\
	\hline
  \end{tabular}
\label{tab_avg}
\end{table}

@@

Regarding the proposed algorithm, \malg, it has been obtained an upload bandwidth usage reduction of 63.51\% and 33.58\%, compared to the circular area of interest and the field of view based algorithms, respectively. The maximum upload bandwidth usage was also reduced, decreasing 52.03\% and 33.10\%, compared to the same algorithms. Table \ref{tab_summary} shows the maximum and average bandwidth saving percentage of the \malg\ algorithm compared to the other techniques. For example, the maximum bandwidth utilization with \malg\ is 24.81\% less than with a circular area of interest based algorithm with update frequency attenuation.

@@

It was also observed that the average and maximum observed values differ, even when there is no algorithm used to manage interest, that is, the client receives state updates of all entities in the game, with maximum frequency. With 200 avatars in the environment, 100 bytes state update packet size and 250 ms normal update interval, the server should be using 80000 bytes per second - 200 $ \ times $ 100 $ \ times $ 4 - for each client. However, it was observed that the maximum and average usage, with 200 avatars present and no interest management, was 77600 and 76458, respectively. This happens because ns-2 is a discrete event simulator, and the simulated server was programmed to check the update schedule every 10 ms. As a result of this, each state update may have had its interval increased by up to 10 ms, which explains the values found.

\section{Conclusion}
\label{conclusion}
In this work, we presented the \malg\ interest management algorithm, whose main idea is to adapt the state update frequency of the game entities according to their relevance to each client that will receive the updates. The \malg\ area of interest consists of a circular sector, corresponding to the player field of view, plus a smaller circle, which represents the area close to that player's avatar. The goal of this close area is to maintain the game state inside it the most up-to-date possible, in order to improve the game near the avatar. Joining these characteristics, we obtained an algorithm that achieved a significant reduction of the maximum upload bandwidth utilization by the server, which decreased 52.03\% and 33.10\%, compared to circle and field of view based interest management algorithms, respectively. The average upload bandwidth utilization by the server also decreased 63.51\% and 33.58\%, compared to the same algorithms.

>>>>

Before sending the state of an entity to a client, the last transmission time is checked. The next transmission is then scheduled to occur after a certain interval. If the relevance of that state is 1, the normal update interval will be used. If it is less than 1, the normal interval is divided by the relevance. For example, let the normal update interval of a game be \mbox{200 ms}. If the avatar $A_i$, which just sent a state update packet to $A_j$, is at a distance from $A_i$ such that its relevance is 0.5, the next transmission will only happen after an interval of 200 divided by 0.5. So, the player controlling $A_j$ will only receive an update of $A_i$ after 400 ms. Despite this interval is still less than a half second, it represents a reduction of the state update frequecy of $A_i$ in 50\%. As they are at a greater distance from one another, and the interval was increased only by 200 ms, this variation will probably be imperceptible by the player who controls $A_j$.

@@

In order to perform the simulation of the proposed algorithm, it was first necessary to create a simulated virtual environment, with avatars present on it, since the algorithm is based on locality and view angle information. The environment consists of a two-dimensional space, which corresponds to the region managed by one of the servers. There are various avatars present, whose number varies from one simulation to another. Each one of them randomly chooses a destination point in the environment and then moves there. When it reaches its destination, it stays for a random time value, which may be zero, and then chooses a new destination to move to.

Ns-2 simulator \cite{mccanne:nsn} was used to simulate and compare the interest management algorithms described in this paper. This simulator allows the user to create code of the specific application which will be simulated. In our case, it was simulated a server managing a certain region where there was an avatar whose client should receive state updates of the other avatars present in that region.
%In our case, it was simulated a server which should send state updates to the client responsible for one of the avatars in the region managed by that server.
Based on their locations and on the chosen interest management algorithm, the server decided which other avatars had a relevant state to the client in question.

Through the simulations, it was obtained the server upload bandwidth usage for one client. It was not considered necessary to simulate every client connected to the server simultaneously because every one of them had the same behavior. To find the total upload bandwidth usage, we just multiply the value found in the simulation by the number of clients connected to that server. Also, there was no point in simulating or measuring the download bandwidth usage of the server, since it would keep receiving the players' actions at the same rate no matter what interest management algorithm was used.

%Another question is that the upload bandwidth usage by the server is much higher than the download one. If it receives $n$ actions, each from one of the $n$ clients connected to it, it must send, in the worst case, $O(n^2)$ state updates, because each player needs the state of the others. Therefore, it was only necessary to measure the upload traffic from the server.

@@

Some works, such as \cite{yu2007nas}, \cite{kim2005tcm} and \cite{svoboda2007taa}, analyse the network traffic generated by large scale games. Based on these analysis, and adopting a conservative posture, the following parameters have been decided to be used in the simulations:

\begin{itemize}
 \item Normal update interval: 250 ms;
 \item State update for one entity: 100 bytes UDP packet;
 \item Duration of each simulated game session: 20 min;
 \item Virtual environment area: 750 x 750;
 \item View distance: 120;
 \item Critical distance: 40;
 \item View angle: 180°.
\end{itemize}

Several simulations were executed, in order to compare the described interest management algorithms. The number of avatars in the environment was varied in order to measure the scalability. The algorithms compared were the ones with circle, circle with attenuation and field of view based area of interest and the proposed algorithm, \malg. To demonstrate how much traffic reduction each one of these achieve, simulations in which no type of interest management was used - and the server sent to the client state updates of all entities in the environment - were also performed.

@@

The results were collected as follows: to measure the average upload bandwidth usage by the server, the sizes of all packets sent in the session were summed up and then divided by the session duration; to determine the maximum usage, it was measured how many bytes had been sent each second and the highest value was selected.

@@

\begin{table}[ht]
\caption{Maximum bandwidth utilization}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9400	&	8500	&	5700	&	7100	&	4700 \\
		50	&	19300	&	17000	&	10300	&	12300	&	8100 \\
		75	&	29100	&	23600	&	16600	&	17800	&	11300 \\
		100	&	38800	&	32500	&	20500	&	23000	&	15500 \\
		125	&	48600	&	37400	&	24300	&	29500	&	19700 \\
		150	&	58300	&	47400	&	29900	&	32900	&	22700 \\
		175	&	67700	&	56100	&	34300	&	32400	&	21500 \\
		200	&	77600	&	62300	&	37500	&	41200	&	28900 \\
	\hline
  \end{tabular}
\label{tab_max}
\end{table}

\begin{table}[ht]
\caption{Average bandwidth utilization}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9221	&	4715	&	2759	&	2534	&	1700 \\
		50	&	18826	&	9350	&	5442	&	4949	&	3303 \\
		75	&	28432	&	13963	&	8315	&	7619	&	5137 \\
		100	&	38037	&	19324	&	11029	&	9928	&	6739 \\
		125	&	47642	&	23138	&	13871	&	12434	&	8290 \\
		150	&	57247	&	29031	&	16432	&	15085	&	10062 \\
		175	&	66853	&	34697	&	19661	&	23060	&	14250 \\
		200	&	76458	& 	38600	&	23450	&	21491	&	14413 \\
	\hline
  \end{tabular}
\label{tab_avg}
\end{table}

@@

Regarding the proposed algorithm, \malg, it has been obtained an average upload bandwidth usage reduction of 63.51\% and 33.58\%, compared to the circular area of interest and the field of view based algorithms, respectively. The maximum upload bandwidth usage was also reduced, decreasing 52.03\% and 33.10\%, compared to the same algorithms. Table \ref{tab_summary} shows the maximum and average bandwidth saving percentage of the \malg\ algorithm compared to the other techniques, for average and maximum usage. For example, the maximum bandwidth utilization with \malg\ is 24.81\% less than with a circular area of interest based algorithm with update frequency attenuation.

@@

It was also observed that the average and maximum usage values differ, even when no algorithm is used to manage interest, that is, the client receives state updates of all entities in the game, with maximum frequency. With 200 avatars in the environment, 100 bytes state update packet size and 250 ms normal update interval, the server should send 80000 bytes per second - 200$\times$100$\times$4 - to each client. However, it was observed that the maximum and average usage, with 200 avatars present and no interest management, was 77600 and 76458, respectively. It happens because \mbox{ns-2} is a discrete event simulator, and the simulated server was programmed to check the update schedule every 10 ms. As a result of this, each state update may have had its interval increased by up to 10 ms, which explains the values found.

\section{Conclusion}
\label{conclusion}
In this work, we presented the \malg\ interest management algorithm, whose main idea is to vary the state update frequency of the game entities according to their relevance to each client that will receive the updates. The \malg\ area of interest consists of a circular sector, corresponding to the player field of view, plus a smaller circle, which represents the area close to that player's avatar. The goal of this close area is to maintain the game state inside it the most up-to-date possible, in order to improve the game near the avatar. Joining these characteristics, we obtained an algorithm that achieved a significant reduction of the maximum upload bandwidth utilization by the server, which decreased 52.03\% and 33.10\%, compared to circle and field of view based interest management algorithms, respectively. The average upload bandwidth utilization by the server also decreased 63.51\% and 33.58\%, compared to the same algorithms.
